### 2012-08-26 mindreading/conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
espp talk done

Can we reject (\ref{infant_perspectives}), the claim that infants neither process perspective differences nor engage in counterfactual reasoning until they are at least one year old?
One difficulty is that current developmental evidence almost uniformly supports this claim  \citep{rafetseder:2010_counterfactual,beck:2011_almost}.
A second problem is that the development of abilities to reason counterfactually,%
\footnote{Here and throughout I follow Eva Rafetseder in using the term `counterfactual reasoning'  to refer to reasoning which essentially involves comparing two possible situations; the term is sometimes used more broadly to include the sorts of representations involved in pretence.} 
like the development of abilities to represent false belief,
appears to involve working memory and inhibitory control 
(on counterfactuals: 
	\citealp{drayton:2011_counterfactual,beck:2011_supporting};
on false beliefs: \citealp{en_1412, en_1698}; \citealp{lin:2010_reflexively, en_1547} experiments 4-5; \citealp{saxe_reading_2006}).
As even committed nativists are likely to agree, capacities for inhibitory control and working memory develop over several years and are limited in infants \citep[e.g.][]{carlson:2005_developmentally}.
So we cannot straightforwardly reject (\ref{infant_perspectives}).

@@

A less obviously flawed approach to rejecting (\ref{fb_is_perspectives}) involves analysing particular false belief tasks.
Two explore this idea it will be useful to distinguish between two types of false belief task, call them Category A and Category B.
By stipulation Category A tasks have these features:
\begin{itemize}
\item Children tend to pass them some time after their third birthday.
\item Abilities to pass these tasks has a protracted developmental course stretching over months if not years.
\item Success on these tasks is correlated with developments in executive function \citep[]{en_410, en_1130} and language \citep[]{en_1209}.
\item Success on these tasks is facilitated by explicit training \citep[]{en_85} and environmental factors such as siblings \citep[]{en_507, en_1299}.  
\item Abilities to succeed on these tasks typically emerge from extensive participation in social interactions (as \citealp{en_1300} suggest).
\end{itemize}
All the other false belief tasks are in Category B.
It is these Category B tasks on which infants succeed.
The proposal we are considering (and which I shall reject) is this.  
\begin{quote}
	All Category A tasks impose a requirement (or set of requirements) other than the requirement to represent a false belief;
call this the \emph{Extraneous Requirement}.
	The connections between success on Category A tasks and processing perspective differences and counterfactual reasoning are a consequence just of the extraneous requirement.
\end{quote}
I am going to argue that this proposal should be rejected.
First note that the proposal is harder to defend than sometimes assumed.
For instance, philosophers sometimes mistakenly assume that all Category A tasks involve language or communication, whereas Category B tasks do not.
But some Category B tasks involve language and communication 
\citep{Knudsen:2011fk,Song:2008qo}
and there are non-verbal tasks in Category A
(\citealp{Call:1999co}; \citealp{low:2010_preschoolers} Study 2).
We should also note that children who fail Category A tasks can answer questions about perception or pretence that are word-for-word identical with the questions about false belief that they cannot answer correctly (\citealp{Gopnik:1994cb}; see also \citealp{Cluster:1996ht}).
So the Extraneous Requirement cannot be straightforwardly linked to language or communication.
There is a more general problem facing attempts to identify the Extraneous Requirement.
There is much variation among the tasks in Category A.
It is not just that some are verbal whereas others are nonverbal.
It is also that some involve prediction whereas others involve retrodiction or justification \citep[e.g.][]{Wimmer:1998kx},
some concern the first-person perspective, whereas others involve a second- or third-person perspective \citep[e.g.][]{Gopnik:1991db},
some involve interaction whereas in others the subject is a mere observer \citep[e.g.][]{Chandler:1989qa},
and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
Despite all this variation and more, the Category A tasks all appear to measure a single developmental transition \citep{Wellman:2001lz}.
This is a significant obstacle to identifying the Extraneous Requirement.
In fact, in my view the currently available evidence supports the conclusion that there is no Extraneous Requirement.
And as far as I know the only plausible candidate for the transition  Category A tasks measure is a transition in the ability to represent beliefs.
This is why I don't think that we can straightforwardly reject  (\ref{fb_is_perspectives}).

@@

\item[3$'$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
\end{description}
\end{quote}
%
When the primary measure of early false belief understanding was  looking times or anticipatory eye movements, this seemed promising \citep{Clements:1994cw,Clements:2000nc,Garnham:2001ql,Ruffman:2001ng}.
However, as already mentioned, we now know that one-year-old infants succeed on false belief tasks which involve actively helping others, interpreting their utterances and pointing to provide information \citep{en_1783,Knudsen:2011fk,southgate:2010fb},
and that older children still fail some nonverbal false belief tasks \citep{Call:1999co}.
All other things being equal, it would be natural to suppose that success on these sorts of task reflects explicit representations of false belief.
Certainly we lack a theoretically coherent and independently empirically motivated account of the distinction between implicit and explicit representation that associates exactly those tasks infants succeed on (strictly: Category B tasks) with implicit representations of false belief.
So we don't know that distinguishing implicit from explicit representations does enable us to avoid contradiction.

@@

At this point it is perhaps tempting to suppose that we already have everything we need to resolve the puzzle.
The idea would be this:
%
\begin{quote}
\begin{description}
\item[1$''$.] Infants can represent false beliefs \emph{using a comparatively simple measure} from around their first birthday or earlier.
%
\item[2$''$.] Being able to represent false beliefs \emph{using a comparatively sophisticated measure}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both). 
%
\item[3$''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  
\end{description}
\end{quote}
%
I think all these claims are true, and with Ian Apperly I have developed in detail a conjecture about the comparatively simple measure infants and others may use in representing beliefs \citep{butterfill_minimal}.
It turns out that different measures of belief generate testable and novel predictions about the limits of subjects' abilities to track beliefs, so although we haven't tested the conjecture yet it is a testable conjecture.

@@

So here is a final refinement of the claims which make up the puzzle that I started with:
%
\begin{quote}
\begin{description}
\item[1$'''$.] Infants can represent false beliefs \emph{using a comparatively simple measure} and \emph{in a modular process} from around their first birthday or earlier.
%
\item[2$'''$.] Being able to represent false beliefs \emph{using a comparatively sophisticated measure} and  \emph{in a non-modular process}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both). 
%
\item[3$'''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  
\end{description}
\end{quote}
%
At this point the claims are no longer contradictory.
But it remains an open question whether we now have all the ingredients necessary to explain both early success on Category B false belief tasks and late failure on Category A tasks or whether  some further complexity in the notion of representing belief also needs to be added to the mix.\footnote{
\citet[p.\ 155-6]{Apperly:2010kx} appears to suggest a solution additionally involving egocentric bias, but I'm not yet fully convinced.
}

>>>>

Can we reject (\ref{infant_perspectives}), the claim that infants neither process perspective differences nor engage in counterfactual reasoning until they well over one year old?
One difficulty is that current developmental evidence almost uniformly supports this claim  \citep{rafetseder:2010_counterfactual,beck:2011_almost}.
A second problem is that the development of abilities to reason counterfactually,%
\footnote{Here and throughout I follow Eva Rafetseder in using the term `counterfactual reasoning'  to refer to reasoning which essentially involves comparing two possible situations; the term is sometimes used more broadly to include the sorts of representations involved in pretence.} 
like the development of abilities to represent false belief,
appears to involve working memory and inhibitory control 
(on counterfactuals: 
	\citealp{drayton:2011_counterfactual,beck:2011_supporting};
on false beliefs: \citealp{en_1412, en_1698}; \citealp{lin:2010_reflexively, en_1547} experiments 4-5; \citealp{saxe_reading_2006}).
As even committed nativists are likely to agree, capacities for inhibitory control and working memory develop over several years and are limited in infants \citep[e.g.][]{carlson:2005_developmentally}.
So we cannot straightforwardly reject (\ref{infant_perspectives}).

@@

A less obviously flawed approach to rejecting (\ref{fb_is_perspectives}) involves analysing particular false belief tasks.
Two explore this idea it will be useful to distinguish between two types of false belief task, call them Category A and Category B.
By stipulation Category A tasks have these features:
\begin{itemize}
\item Children tend to pass them some time after their third birthday.
\item Abilities to pass these tasks has a protracted developmental course stretching over months if not years.
\item Success on these tasks is correlated with developments in executive function \citep[]{en_410, en_1130} and language \citep[]{en_1209}.
\item Success on these tasks is facilitated by explicit training \citep[]{en_85} and environmental factors such as siblings \citep[]{en_507, en_1299}.  
\item Abilities to succeed on these tasks typically emerge from extensive participation in social interactions (as \citealp{en_1300} suggest).
\end{itemize}
All the other false belief tasks are in Category B.
It is these Category B tasks on which infants succeed.
The proposal we are considering (and which I shall reject) is this.  
\begin{quote}
	All Category A tasks impose a requirement (or set of requirements) other than the requirement to represent a false belief;
call this the \emph{Extraneous Requirement}.
	The connections between success on Category A tasks and processing perspective differences and counterfactual reasoning are a consequence just of the extraneous requirement.
\end{quote}
I am going to argue that this proposal should be rejected.
First note that the proposal is harder to defend than sometimes assumed.
For instance, philosophers sometimes mistakenly assume that all Category A tasks involve language or communication, whereas Category B tasks do not.
But some Category B tasks involve language and communication 
\citep{Knudsen:2011fk,Song:2008qo}
and there are non-verbal tasks in Category A
(\citealp{Call:1999co}; \citealp{low:2010_preschoolers} Study 2).
We should also note that children who fail Category A tasks can answer questions about perception or pretence that are word-for-word identical with the questions about false belief that they cannot answer correctly (\citealp{Gopnik:1994cb}; see also \citealp{Cluster:1996ht}).
So the Extraneous Requirement cannot be straightforwardly linked to language or communication.
There is a more general problem facing attempts to identify the Extraneous Requirement.
There is much variation among the tasks in Category A.
It is not just that some are verbal whereas others are nonverbal.
It is also that some involve prediction whereas others involve retrodiction or justification \citep[e.g.][]{Wimmer:1998kx},
some concern the first-person perspective, whereas others involve a second- or third-person perspective \citep[e.g.][]{Gopnik:1991db},
some involve interaction whereas in others the subject is a mere observer \citep[e.g.][]{Chandler:1989qa},
and some involve predicting actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
Despite all this variation and more, the Category A tasks all appear to measure a single developmental transition \citep{Wellman:2001lz}.
This is a significant obstacle to identifying the Extraneous Requirement.
In fact, in my view the currently available evidence supports the conclusion that there is no Extraneous Requirement.
And as far as I know the only plausible candidate for the transition  Category A tasks measure is a transition in the ability to represent beliefs.
This is why I don't think that we can straightforwardly reject  (\ref{fb_is_perspectives}).

@@

\item[3$'$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  \label{infant_perspectives}
\end{description}
\end{quote}
%
When the primary measure of early false belief understanding was  looking times or anticipatory eye movements, this seemed promising \citep{Clements:1994cw,Clements:2000nc,Garnham:2001ql,Ruffman:2001ng}.
However, as already mentioned, we now know that one-year-old infants succeed on false belief tasks which involve actively helping others, interpreting their utterances and pointing to provide information \citep{en_1783,Knudsen:2011fk,southgate:2010fb},
and that older children still fail some nonverbal false belief tasks \citep{Call:1999co}.
All other things being equal, it would be natural to suppose that success on these sorts of task reflects explicit representations of false belief.
Certainly we lack a theoretically coherent and independently empirically motivated account of the distinction between implicit and explicit representation that associates exactly those tasks infants succeed on (strictly: Category B tasks) with implicit representations of false belief.
So we don't know that distinguishing implicit from explicit representations does enable us to avoid contradiction.

@@

At this point it is perhaps tempting to suppose that we already have everything we need to resolve the puzzle.
The idea would be this:
%
\begin{quote}
\begin{description}
\item[1$''$.] Infants can represent false beliefs \emph{using a comparatively simple measure} from around their first birthday or earlier.
%
\item[2$''$.] Being able to represent false beliefs \emph{using a comparatively sophisticated measure}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both). 
%
\item[3$''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  
\end{description}
\end{quote}
%
I think all these claims are true, and with Ian Apperly I have developed in detail a conjecture about the comparatively simple measure infants and others may use in representing beliefs \citep{butterfill_minimal}.
It turns out that different measures of belief generate testable and novel predictions about the limits of subjects' abilities to track beliefs, so although we haven't tested the conjecture yet it is a testable conjecture.

@@

So here is a final refinement of the claims which make up the puzzle that I started with:
%
\begin{quote}
\begin{description}
\item[1$'''$.] Infants can represent false beliefs \emph{using a comparatively simple measure} and \emph{in a modular process} from around their first birthday or earlier.
%
\item[2$'''$.] Being able to represent false beliefs \emph{using a comparatively sophisticated measure} and  \emph{in a non-modular process}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both). 
%
\item[3$'''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  
\end{description}
\end{quote}
%
At this point the claims are no longer contradictory.
But it remains an open question whether we now have all the ingredients necessary to explain both early success on Category B false belief tasks and late failure on Category A tasks or whether  some further complexity in the notion of representing belief also needs to be added to the mix.\footnote{
\citet[p.\ 155-6]{Apperly:2010kx} appears to suggest a solution additionally involving egocentric bias, but I'm not yet fully convinced.
}
